Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-06-30 article-models-ttinternal.tex
improvement identity terms
attempting to write the rule for cftidtm
finally stated the rule that weakening is compatible with itself

\subsubsection{Weakening preserves itself}
The following diagram displays the compatibility of weakening with itself:
\begin{equation*}\label{diag:cftwk-cftwk}
\begin{tikzcd}[column sep=huge]
\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
& \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
  \ar{d}{\jcomp{}{\cftctxext}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
  \\
\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_A}}}}}
  \ar{r}[swap]{\jcomp{}{\cftwkc{A}}{\jcomp{}{\cftctxext}{\cftwk{A}}}}
& \jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_{F_A}}}}}
\end{tikzcd}
\end{equation*}
This is a complicated diagram and it is not immediately clear that all of it
makes sense. Therefore we examine its ingredients in more detail.

@@

\begin{lem}
Let $\cftalg{A}$ be a term-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a homomorphism of term-algebras satisfying \autoref{diag:cftwk-cftwk}. Then
we have the judgmental equality
\begin{equation*}
\jtermeq
  {}
  {}
  {}
  {}
\end{equation*}
for $\gamma:A$, $x:\cftalgf{\cftalg{A}}$, $y_0:\cftalgf{\cftalg{A}}$ and
$y_1:\cftalgf{\cftalg{F_A}}$.
\end{lem}

@@

\subsection{Projection algebras}
Projection algebras will be weakening algebras with additional terms implementing
the identity terms. We call these algebras projection algebras because the
identity terms can only be formulated in weakening algebras and together with
weakening, the identity terms provide for all the projections. In a weakening
algebra $\cftalg{A}$ in context $\Gamma$, identity terms are implemented by a term
\begin{equation*}
\jterm
  {context}
  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftalg{F_{F_A}}}}}
  {\cftidtm{\cftalg{A}}}
\end{equation*}
We will require additionally that identity terms are compatible with weakening.
Thus we will formulate an abstraction of the rule stated in \autoref{comp-wi}.

\subsection{Substitution algebras}
Substitution algebras will be CFT-algebras with an additional substitution 
operation. Given a CFT-algebra $\cftalg{A}$ in context $\Gamma$, substitution 
will be a CFT-homomorphism
\begin{equation*}
\jhom
  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
  {\cftalgf{\cftalg{A}}}
  {\cftsubst{\cftalg{A}}}
\end{equation*}
We will require additionally that substitution preserves itself. Thus, we will
formulate an abstraction of the rules stated in \autoref{comp-ss}.

>>>>

\subsubsection{Weakening preserves itself}
The following diagram displays the compatibility of weakening with itself:
\begin{equation}\label{diag:cftwk-cftwk}
\begin{tikzcd}[column sep=10em]
\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
& \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
  \\
\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_A}}}}}
  \ar{r}[swap]{%
    \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftwk{\cftalg{A}}}
                }
            }
        }
    }
& \jcomp{}{\cftwkc{A}}{\cftalg{F_{F_{F_{F_A}}}}}
\end{tikzcd}
\end{equation}
This is a complicated diagram and it is not immediately clear that all of it
makes sense. Therefore we examine its ingredients in more detail.

@@

\begin{lem}
Let $\cftalg{A}$ be a term-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a homomorphism of term-algebras for which the diagram in
\autoref{diag:cftwk-cftwk} commutes judgmentally. Then
we have the judgmental equality
\begin{align*}
& \jcomp{}
  { \subst{q}{{{b}{{\gamma}{\cftctxext}}}{\cftwk{\cftalg{A}}}}
    }
  { \subst
      {{q}{{b}{\cftfamext}}}
      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
    }
  \\
& \jdeq
  \jcomp{}
    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
      }
    { \subst
        { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
        { {{{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}{{{a}{{\gamma}{\cftctxext}}}{\cftctxext}}}
          {\cftwk{\cftalg{A}}}
          }
      }
\end{align*}
of CFT-homomorphisms from $\subst{b}{{\gamma}{\cftalg{F_{F_A}}}}$ to
$ \subst
      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}
        }
      { {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
        {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
        }$
in context $\Gamma$, for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$q:\subst{b}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$.
\end{lem}

@@

\subsection{Projection algebras}
Projection algebras will be weakening algebras with additional terms implementing
the identity terms. We call these algebras projection algebras because the
identity terms can only be formulated in weakening algebras and together with
weakening, the identity terms provide for all the projections. In a weakening
algebra $\cftalg{A}$ in context $\Gamma$, identity terms are implemented by a term
\begin{equation*}
\jterm
  { { {\Gamma}
      {\cftalgc{\cftalg{A}}}
      }
    { \cftalgf{\cftalg{A}}
      }
    }
  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftalg{F_A}}}}
  {\cftidtm{\cftalg{A}}}
\end{equation*}
We will require additionally that identity terms are compatible with weakening.
Thus we will formulate an abstraction of the rule stated in \autoref{comp-wi}.

\begin{equation}
\jtermeq
  { \xi}
  { \xi}
  { \subst
      { \jcomp
          { }
          { \cftctxext}
          { \cftidtm{\cftalg{A}}
            }
        }
      { { { \idtm
              { \jcomp
                  { }
                  { \cftctxext}
                  { \cftalgf{\cftalg{A}}
                    }
                }
            }
            { \jcomp
                { }
                { \cftctxext
                  }
                { \cftwkc{\cftalg{A}}
                  }
              }
          }
        { \jcomp
            { }
            { \jcomp{}{\cftctxext}{\cftfamext}}
            { \cftwkt{\cftalg{A}}}
          }
        }
    }
  { \xi}
\end{equation}

\subsection{Substitution algebras}
Substitution algebras will be CFT-algebras with an additional substitution 
operation. Given a CFT-algebra $\cftalg{A}$ in context $\Gamma$, substitution 
will be a CFT-homomorphism
\begin{equation*}
\jhom
  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
  {\cftalg{F_A}}
  {\cftsubst{\cftalg{A}}}
\end{equation*}
We will require additionally that substitution preserves itself. Thus, we will
formulate an abstraction of the rules stated in \autoref{comp-ss}.
